---
title: "Blocks"
output: html_document
---

Search for differentially methylated blocks
```{r}
library(minfi)
path="~/25Apr/tcgaMethylationSubset-master"
targets=read.delim(file.path (path,"targets.txt"),as.is=TRUE)
index = which( targets$Tissue=="colon")
targets = targets[index,]
dat = read.450k.exp(base=path,targets = targets, verbose=TRUE)
```

convert raw data into an object that includes methylation values, as well as the location of CpGs

```{r}
library(IlluminaHumanMethylation450kmanifest)
library(IlluminaHumanMethylation450kanno.ilmn12.hg19)
dat <- preprocessIllumina(dat)
dat <- mapToGenome(dat)
```

collapse the data

```{r}
cdat <- cpgCollapse(dat)
class(cdat)
```

How many regions are represented in the collapsed object?
```{r}
cdat$object
```

We can see the type of regions that are represented in this collapsed object in the type column:
```{r}
head(granges(cdat$object))
```

What proportion of the regions are OpenSea regions?
```{r}
prop.table(table(mcols(granges(cdat$obj))$type))
```

use the blockFinder function to find differentially methylated regions between cancer and normal:
```{r}
status <- factor(pData(cdat$object)$Status, level=c("normal", "cancer"))
X <- model.matrix(~status)
res <- blockFinder(cdat$object, X, cutoff=.05)
```

blockFinder calls bumphunter and returns a similar object.
```{r}
head(res$table)
```

What proportion of the blocks reported in res$table are hypomethyated (lower methylation in cancer versus normal)?
```{r}
mean(res$table$value < 0)

tab <- makeGRangesFromDataFrame(res$table)
index <- granges(cdat$object) %over% (tab[1]+10000)
pos <- start(cdat$object)[index]
col <- as.numeric(status)
matplot(pos, getBeta(cdat$object)[index, ], col=col, pch=1, cex=.5)

plot(pos, res$fitted[index])
```

