---
title: "Fun with genome 1"
output: html_document
---

From [My blogger fun-with-genome-1](http://bitit1994.blogspot.hk/2016/04/fun-with-genome-1.html)

It's final exam's week. Every one is buried in their study. I'm now taking a break and making some fun with the human genome package BSgenome.Hsapiens.UCSC.hg19. 

This package contains the entire genome sequences for human provided by University of California, Santa Cruz (UCSC).

```{r message=FALSE, warning=FALSE}
library(BSgenome.Hsapiens.UCSC.hg19)
```

There is a set of generic functions for us to get or modify the sequence information like seqinfo, seqnames, seqlengths...

```{r}
genome <- BSgenome.Hsapiens.UCSC.hg19
seqinfo(genome)
seqnames(genome)
seqlengths(genome)
```

It is in April now, so I will take chromosome 4 for example.

```{r}
chr4 <- genome$chr4
```

The DNA string is still too large. Let me cut it down.

```{r}
s <- subseq(chr4, start=20160425, width=999)
```

I'm curious about What it looks like.

```{r}
print(s)
print(as.character(s))
```

Emm.. I can see nothing but four letters. 
I should ask more specifically.
What is the GC-content of my new sequence? 

```{r}
sum(alphabetFrequency(s)[c("C","G")])
```

How about a proportion?
```{r}
letterFrequency(s,"CG",as.prob=TRUE)
```

Here is a function I find useful. countPattern

```{r}
countPattern("CG",s)
```

It answers the question: how many times a given pattern occurs in a reference sequence.

I can also locate them
```{r}
matchPattern("CG",s)
```
