---
title: "The t-test for CpG islands and volcano plot"
output: html_document
---

From [My blogger The t-test for CpG islands and volcano plot](http://bitit1994.blogspot.hk/2016/04/the-t-test-for-cpg-islands-and-volcano.html)

In DNA methylation data analysis, t-test statistics is used to identify differences in DNA methylation at single CpG sites. 

Now I'm going to use a modified t-test in an R package, limma, to identify differentially methylated CpGs between samples from colon cancer and normal tissues.

```{r warning=FALSE}
library(limma)
```

1. Load data.

```{r}
load("dna1.rda")
```

2. Baseline model design matrix

The design matrix indicates which arrays are from  cancer tissues.

```{r message=FALSE, warning=FALSE}
design <- model.matrix(~pd$Status)
```

3. Fit the baseline model

Fit a linear model for each gene to estimate the fold changes and standard errors.

```{r}
fit <- lmFit(meth, design)
```

4. empirical Bayes moderation

Apply empirical Bayes smoothing to the standard errors.

```{r}
eb <- ebayes(fit)
```

A volcano plot reveals effect size on x-axis and the statistical significance on y-axis so that highly dysregulated genes appear farther to the right and left sides while highly significant changes are higher on the plot.

```{r warning=FALSE}
library(ggplot2)

fc = fit$coef[,2]
sig = -log10(eb$p.value[,2])
df <- data.frame(fc, sig)
df$thre <- as.factor(abs(fc) < 0.4 & sig < -log10(0.05))

ggplot(data=df, aes(x=fc, y = sig, color=thre)) +
  geom_point(alpha=.6, size=1.2) +
  theme(legend.position="none") +
  xlab("Effect size") +
  ylab("-log10 p value")

```

Why do these two colors match so well? I like it!